---
title: "test"
author: "Zachary Richardson"
date: "`r Sys.Date()`"
output: html_document
---

## Description
The current goal of this project is to identify areas of the court where assists are most impactful. This can be useful for identifying areas on the court to design plays resulting in catch-and-shoot opportunities. This can also be used to identify which players excel more in specific areas of the court off of assisted shots, as well as who is not impact (or potentially negatively impacted) by assisted shot opportunities.

To complete this analysis, we need several points of data:
1. Shot location (x, y)
2. Shot outcome (Make/Miss)
3. Shot assisted (Y/N)

There exists a data set in GitHub uploaded by DomSamangy which contains shot data for every shot taken from 2003-2024 seasons. It does not contain assist information, so we need to scrape play-by-play data from Basketball Reference, then merge the data sets using time stamps to determine whether a shot was assisted or not. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE}
## Load packages
library(tidyverse)
library(rvest)
library(stringr)
library(magrittr)
library(lubridate)

## Ensure dplyr verbs work correctly
mutate <- dplyr::mutate
select <- dplyr::select
filter <- dplyr::filter
arrange <- dplyr::arrange
summarize <- dplyr::summarize
```


```{r}
## Import shots data (credit to DomSamangy on GitHub for the data set)
shots = read.csv("NBA_2025_Shots.csv")
```

## Web Scraping

```{r}
## Test of rvest to web scrape a box score from a specific game on Basketball Reference
#raw_pbp = read_html("https://www.basketball-reference.com/boxscores/pbp/202410220BOS.html")



## Define a function that takes a raw_html (play-by-play table) and returns a tidy/clean data frame
tidy_pbp <- function(raw_pbp){
  
  ## Use selector gadget tool in chrome to identify the html elements to scrape
  
  ## Scrap game data and time info
  game_date_time = raw_pbp %>%
    html_elements(".scorebox_meta div:nth-child(1)") %>%
    html_text() # keep only text
  
  ## Scrape away and home team names
  teams = raw_pbp %>%
    html_elements(".scorebox strong a") %>%
    html_text() # keep only text
  
  
  ## Scrape and clean/tidy play-by-play data
  pbp_clean = raw_pbp %>%
    html_table() %>%
    pluck(1) %>%
    
    ## Change column names
    set_names(c("time", "away_detail", "away_result", "score", "home_result", "home_detail")) %>%
    
    ## Add in period column (allow for up to 4 OT since this has happened semi-recently)
    mutate(period = case_when(
      score == "End of 1st quarter" ~ 1,
      time == "2nd Q" ~ 2,
      time == "3rd Q" ~ 3,
      time == "4th Q" ~ 4,
      time == "1st OT" ~ 5,
      time == "2nd OT" ~ 6,
      time == "3rd OT" ~ 7,
      time == "4th OT" ~ 8
    )) %>%
    
    ## Fill period information to each row
    fill(period, .direction = "downup") %>%
    
    ## Remove data-less rows
    filter(score != "Score",
           !str_detect(score, "End"), # Rows indicating end of quarter 
           !if_any(c(away_detail, home_detail), ~ str_detect(.x, regex("shot clock")))) %>% # Shot clock violations
    
    ## Keep only rows containing shot information (removes free throws as well)
    filter(if_any(c(away_detail, home_detail), ~ str_detect(.x, regex("shot | layup | dunk")))) %>%
    
    ## Convert _result columns to numerical
    mutate(across(c(away_result, home_result), ~ as.numeric(gsub("\\+", "", .))),
           across(c(away_result, home_result), ~ ifelse(is.na(.), 0, .)),
           away_team = teams[1],
           home_team = teams[2],
           game_info = game_date_time) %>%
    
    ## Split time column into minutes and seconds columns
    separate_wider_delim(time, ":", names = c("min_left", "sec_left")) %>%
    separate_wider_delim(score, "-", names = c("away_score", "home_score")) %>%
    mutate(across(c(min_left, sec_left, away_score, home_score), ~ as.numeric(.))) %>%
    
    ## Split game_info into date and time columns
    ## Use too_many = "merge" to keep date together and only split at the first comma
    separate_wider_delim(game_info, ",", names = c("start_time", "game_date"), too_many = "merge") %>%
    mutate(game_date = gsub("\\,", "", game_date), # Remove comma from the date
           game_date = gsub("^ ", "", game_date), # Remove space from the beginning of the date
           game_date = as.Date(game_date, format = "%B %d %Y")) # Convert into a date in YYYY-MM-DD Format

  return(pbp_clean)
  
}

```


```{r}
## TODO: Find a way to scrape every game from every day, month, year since ?


## Create a function that loops through daily boxscores given a desired range of dates

scrape_pbp <- function(start_date, end_date){
  
  pbp_stats_list = list()
  
  ## Loop through each date in the given date range
  for(date in start_date:end_date){
    
    print(paste0("Today's date is ", as.Date(date)))
    
    ## Pull day, month, and year values
    day_str = as.Date(date) %>% day()
    month_str = as.Date(date) %>% month()
    year_str = as.Date(date) %>% year()
    
    ## Scrape the links to each box score for the given day
    daily_box_score_list = read_html(paste0("https://www.basketball-reference.com/boxscores/?month=", month_str, "&day=", day_str, "&year=", year_str)) %>%
      html_elements(".gamelink") %>%
      html_elements("a") %>%
      html_attr("href") %>%
      gsub("/boxscores/", "", .) # Remove "boxscores" from string due to "pbp" being between it and game id
    
    ## Loop through each game within the daily schedule and scrape play-by-play table
    ## TODO: START HERE: If there are no games on that day, skip to the next date in the for loop
    for(i in 1:length(daily_box_score_list)){
      
      raw_pbp = read_html(paste0("https://www.basketball-reference.com/boxscores/pbp/", daily_box_score_list[i]))
      
      pbp_stats_list[[length(pbp_stats_list) + 1]] = tidy_pbp(raw_pbp)
      
    }
    
    
  }
  
  pbp_df = bind_rows(pbp_stats_list)
  
  return(pbp_df)

}
  
start_date = as.Date("2025-05-14")
end_date = as.Date("2025-05-14")

test = scrape_pbp(start_date, end_date)


```


## Merge Data Sets Together
```{r}
## Example of how to merge once we have all of the games scraped into one data set
## This works, but would make sense to change columns of the shots data set prior to merging 

test = pbp_clean %>%
  mutate(away_team_abbr = "NYK",
         home_team_abbr = "BOS")

shots_test = shots %>%
  filter(GAME_DATE == "10-22-2024") %>%
  mutate(GAME_DATE = as.Date(GAME_DATE, format = "%m-%d-%Y"))


nba_shots_pbp = shots_test %>%
  merge(test,
        by.x = c("GAME_DATE", "AWAY_TEAM", "HOME_TEAM", "QUARTER", "MINS_LEFT", "SECS_LEFT"),
        by.y = c("game_date", "away_team_abbr", "home_team_abbr", "period", "min_left", "sec_left"))


test.FINAL = nba_shots_pbp %>%
  arrange(QUARTER, -MINS_LEFT, -SECS_LEFT) %>%
  select(QUARTER, MINS_LEFT, SECS_LEFT, away_score, home_score, everything())
```








